\chapter*{Problem Description}
\section*{Human Computable Passwords - design and analysis.}
Managing passwords is a significant problem for most people in the modern world. This project will be based around the paper "Human Computable Passwords" by Blocki et al. ~\cite{hcp-blocki}, proposing a method for humans to be able to re-compute passwords from public and reliable storage. Passwords are calculated using a memorized mapping from objects, typically letters or pictures, to digits; the characters of the passwords are then calculated in the users head, using a human computable function.
\par The goal of the project is to determine if the scheme is usable and if it could be used to greater effect in real scenarios. The project will include an implementation of the password management scheme, testing the usability of it and assessing the validity through experiments. The main objectives of the project can be summarized as the following:

\begin{itemize}
    \item Understand and compare the "Human Computable Passwords" scheme with other related password management schemes.
    \item Design and implement a password management scheme applying the ideas of the scheme.
    \item Analyze if the construction could be utilized to provide secure password management in practical situations.
    \item Validate if the scheme is feasible to use, comparing the user efforts required to the security rewards.

\end{itemize}

%to be removed after delivering the project description
\noindent
[1] J. Blocki, M. Blum, and A. Datta, “Human Computable Passwords,” CoRR, vol. abs/1404.0, 2014.

\noindent
\textbf{Assignment given: }12 January, 2015 \\
\textbf{Student: }Anders Kofoed \hfill \textbf{Professor: }Colin Boyd, ITEM \\

